Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-09-27 paper/main.tex
small fix

Parser parser combinators~\cite{vanTonderPPC} is a recent approach enabling
syntax-transformation for multiple languages, and (we show) an apt choice for
mutation testing operators. This key idea is to replace code constructs
common to many languages (e.g., multi-line code blocks delineated by braces)
with simple declarative patterns. Using parser combinators for coarse syntax
matching avoids the complexity of integrating heterogeneous rewrite tools for
mutation testing (e.g., one for Java, one for C++) and human effort to
implement language-specific mutation operators (e.g., where the user has to
learn to write operators for each language tool). The mechanization and
expressive properties of parser parser combinators is covered in prior
literature~\cite{vanTonderPPC}; in this paper we demonstrate that their application is
uniquely effective for universal source-based mutation testing. In
particular, we
build mutation operators with Comby\footnote{https://comby.dev}, the tool that implements parser parser
combinators to declaratively match and rewrite source code syntax.

@@

Comby provides mechanisms to further define custom match syntax and behavior
beyond these standard patterns.\footnote{https://comby.dev/docs/advanced-usage\#custom-language-definitions} Comby patterns can also embed regular expressions, subsuming their
expressive power to enable regex-based mutation operators intermixed with
context-free syntax
properties\footnote{https://comby.dev/docs/basic-usage\#using-regular-expressions}.

@@

As future work, we plan to give users more control over which mutation
operators are used, with controls common to multiple languages.

%We
%add that our approach in general seems to be a good fit to modern
%large software projects, which are often
%multilingual~\cite{Multilingual}: a single mutation tool can be used
%for an entire project.

>>>>

Parser parser combinators~\cite{vanTonderPPC} is a recent approach enabling
syntax-transformation for multiple languages, and (we show) an apt choice for
mutation testing operators. This key idea is to replace code constructs
common to many languages (e.g., multi-line code blocks delineated by braces)
with simple declarative patterns. Using parser combinators for coarse syntax
matching avoids the complexity of integrating heterogeneous rewrite tools for
mutation testing (e.g., one for Java, one for C++) and human effort to
implement language-specific mutation operators (e.g., where the user has to
learn to write operators for each language tool). The mechanization and
expressive properties of parser parser combinators is covered in prior
literature~\cite{vanTonderPPC}; in this paper we demonstrate that their application is
uniquely effective for universal source-based mutation testing. In
particular, we
build mutation operators with Comby (\url{https://comby.dev}), the tool that implements parser parser
combinators to declaratively match and rewrite source code syntax.

@@

Comby provides mechanisms to further define custom match syntax and behavior
beyond these standard patterns. Comby patterns can also embed regular expressions, subsuming their
expressive power to enable regex-based mutation operators intermixed with
context-free syntax
properties.

@@

As future work, we plan to give users more control over which mutation
operators are used, with controls common to multiple languages.
Our approach in general seems to be a good fit to modern
large software projects, which are often
multilingual~\cite{Multilingual}: a single mutation tool can be used
for an entire project.
